Verifying open clusters using the Gaia catalogue

### 2019-01-13 article.tex
Corrections

Using the data from the UCAC4 catalogue, Sampedro et al. reported the existence of 1876 open clusters, using three different membership methods \cite{sampedro}. With the release of the Gaia DR2 catalogue, which contains more precise astrometric data, it is now possible to find and verify the existence of clusters more accurately.
There have already been attempts to identify open clusters using the Gaia DR2 catalogue, for instance, in the study of Cantat-Gaudin et al., 1269 clusters were identified with the new data \cite{cantat-gaudin}. Our first goal will be to crossmatch the data from both studies to find the clusters that haven't already been verified.

\section{The Gaia DR2 catalogue}
The second release of data from Gaia, in the form of the Gaia DR2 catalogue, introduces high-precision measurements of various astrometry parameters, like proper motions and parallaxes, as well as photometric information for a total of $1 692 919 135$ sources \cite{gaiadr2}. The catalogue contains a 5-parameter astrometric solution for over 1.3 billion stars, which include the position in the sky given by the right ascension and declination, the proper motion and the parallax. For the rest of the stars, which are the very faint ones ($G > 21$), a 2-parameter astrometric solution is given instead, with only the positions in the sky.

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=0.45]{ppmxl_vs_gaia_motions}
\caption{Comparison of the data from PPMXL and Gaia DR2 for the proper motions of the cluster Ivanov 2 \cite{ivanov}.}
\label{ppmxl_vs_gaiadr2}
\end{figure}

@@

As a first measure to reduce the number of clusters that need to be verified, we will only check the clusters whose name is not present in the data from both studies. By applying said filter, we end up with a list of 890 clusters, slightly more than expected, due mainly to naming inconsistencies and possibly to one study having found some clusters that the other has not.

For each of these 890 clusters, we compute the mean right ascension $\alpha$ and declination $\delta$, which we will use as the center point of a circle with radius $\dfrac{\max(D) - \min(D)}{2}$ where $D$ is the set of declinations for all the stars of each cluster. Then, we query the data from the Gaia DR2 catalogue for the region encompassed by this circle. We get stars with a magnitude $G < 17$, since the data from the UCAC4 catalogue does not show stars fainter than $G = 17$. This gives us the data for all the stars in the queried region, so we crossmatch this data with the data from the original study to leave only the stars that are listed in Sampedro's study as being part of the cluster. To filter out the repeated cases due to naming inconsistencies, the center positions of the already verified clusters are also added in the position plots.

@@

Figure \ref{crossmatched_data} shows one example of the plots that were made for each of the clusters, for one of the cases that are confirmed to be a cluster and that will be studied in the next section to determine its age and interstellar extinction. Note that, in this case, the proper motion space is pretty compact, and there is a clear straight line in the parallax versus magnitude plot, showing that the stars are moving together and are at the same distance. In the color-magnitude diagram, we can see that some stars have already left the main sequence. We will study this in more detail in the following section.

\section{Determining the age and interstellar extinction of the cluster}
In this section we discuss the procedure followed to determine an approximation of the age and interstellar extinction of some clusters. We try this method for a total of 14 clusters, some of which are taken from the list of verified clusters in the study of Cantat-Gaudin, and the rest are cases from the study of Sampedro which we have verified in the previous section. The complete list of tested clusters can be found in Table \ref{tab:clusters}, along with some parameters.
%TODO comment on the method used to get the members of the cluster
Note that a lower parallax means a greater distance.

\begin{table}[h!]
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Cluster} & \textbf{Position (deg)} & \textbf{Parallax (mas)} & \textbf{Members} \\
\hline
Alessi 3 & $(109.95, -46.05)$ & $3.5619$ & $243$ \\
\hline
Berkeley 100 & $(351.48, 63.78)$ & $0.1461$ & $57$ \\
\hline
Collinder 421 & $(305.83, 41.69)$ & $0.8169$ & $283$ \\
\hline
ESO 332 08 & $(253.80, -40.96)$ & $0.5332$ & $417$ \\
\hline
Herschel 1 & $(116.72, 0.15)$ & $3.3385$ & $149$ \\
\hline
NGC 752 & $(29.24, 37.77)$ & $2.2245$ & $263$ \\
\hline
NGC 1039 & $(40.51, 42.74)$ & $1.9350$ & $659$ \\
\hline
NGC 2126 & $(90.65, 49.91)$ & $0.7490$ & $151$ \\
\hline
NGC 2169 & $(92.12, 13.94)$ & $0.9863$ & $133$ \\
\hline
NGC 2287 & $(101.51, -20.70)$ & $1.3503$ & $673$ \\
\hline
NGC 2360 & $(109.44, -15.63)$ & $0.8994$ & $695$ \\
\hline
NGC 2479 & $(118.76, -17.73)$ & $0.6311$ & $159$ \\
\hline
Ruprecth 8 & $(105.41, -13.56)$ & $0.4278$ & $138$ \\
\hline
Stock 6 & $(35.93, 63.79)$ & $0.9536$ & $144$ \\
\hline
\end{tabular}
\caption{List of clusters that have been studied, including their mean position (given in ($\alpha$, $\delta$)) and mean parallax, computed using the data from Gaia DR2, and the number of member stars of the cluster.} %TODO mention method name
\label{tab:clusters}
\end{table}

@@

Given that the stars in an open cluster are relatively close together, we shift every star in the plot by the distance modulus of the mean parallax of the cluster. We chose this approach instead of using shifting each star by its individual distance modulus because for more distant stars the error in the parallax is higher, so there might be a lot of discrepancy between the parallaxes of some of the stars in the cluster and the CMD would be less clear. Another option would have been to treat the distance modulus as a third parameter to estimate via isochrone fitting, which would also avoid this problem but would make the process harder.

An isochrone is the line that would be drawn in a CMD if we varied the mass of the star at a fixed age. Since the stars in an open cluster have the same age, we can find an isochrone that fits the CMD of the stars in the cluster to determine its age. The isochrones used are from PARSEC \cite{parsec}, with the photometric system from Gaia DR2, and including the interstellar extinction $A_V$ on a star-to-star basis \cite{girardi}. The metallicity value used for the isochrones is $Z = 0.0152$. An example of the result of fitting an isochrone to the cluster data can be seen in Figure \ref{nice-isochrone}.

@@

\section{Problems when fitting isochrones}
\subsection{Unresolved binary stars}
Many stellar systems are binary stars, systems formed by two stars orbiting each other. If they are distant, some these systems may be unresolved and seen as a single object. In a CMD, unresolved binary stars would appear as a secondary line in the main sequence which is brighter than the rest. A clear example of this can be seen in Figure \ref{nice-isochrone}. This can make the main sequence line in the CMD appear wider than it actually is, and can be a problem when fitting isochrones.

@@

\subsection{Differential extinction}
In this work, the reddening due to the interstellar extinction had a fixed value and was included in the isochrones used to estimate the parameters of the clusters. However, in some cases, like for instance Stock 6, the extinction is not the same for all stars in the cluster. In the case of Stock 6, more distant stars appeared to have significantly more extinction, and this makes the stars appear more scattered in the CMD, making the fitting harder. TO avoid this problem, we would have to estimate the extinction for each star individually, and apply the corrections on each of the stars, instead of applying a global extinction for all the stars in the cluster.

>>>>

Using the data from the UCAC4 catalogue, Sampedro et al. reported the existence of 1876 open clusters, using three different membership methods \cite{sampedro}. With the release of the Gaia DR2 catalogue, which contains more precise astrometric data, it is now possible to find and verify the existence of clusters more accurately.
There have already been attempts to identify open clusters using the Gaia DR2 catalogue, for instance, in the study of Cantat-Gaudin et al., 1229 clusters were identified with the new data \cite{cantat-gaudin}.

As a first step, we will crossmatch the data from both studies to find the clusters that haven't already been verified. We will then use Gaia's data to find some cases from the list which are clearly clusters. Finally, we will estimate the age and total interstellar extinction of these clusters.

\section{The Gaia DR2 catalogue}
The second release of data from Gaia, in the form of the Gaia DR2 catalogue, introduces high-precision measurements of various astrometry parameters, like proper motions and parallaxes, as well as photometric information for about 1.7 billion sources \cite{gaiadr2}. The catalogue contains a 5-parameter astrometric solution for over 1.3 billion stars, which include the position in the sky given by the right ascension and declination, the proper motion and the parallax. For the rest of the stars, which are the very faint ones ($G > 21$), a 2-parameter astrometric solution is given instead, with only the positions in the sky.

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=0.45]{ppmxl_vs_gaia_motions}
\caption{Comparison of the data from PPMXL and Gaia DR2 for the proper motions of the cluster Ivanov 2 \cite{ivanov}. We can see that the proper motions also look more compact with Gaia's data, confirming that it is in fact an open cluster.}
\label{ppmxl_vs_gaiadr2}
\end{figure}

@@

As a first measure to reduce the number of clusters that need to be verified, we will only check the clusters whose name is not present in the data from both studies. By applying said filter, we end up with a list of 890 clusters, which more than the 647 we expected, due mainly to naming inconsistencies and possibly to one study having found some clusters that the other has not. However, our main focus will be the 647 clusters that Sampedro found with the less precise data from UCAC4.

For each of these 890 clusters, we compute the mean right ascension $\alpha$ and declination $\delta$, which we will use as the center point of a circle with radius $\dfrac{\max(\delta) - \min(\delta)}{2}$, where we consider the set of declinations of the stars in each cluster. Then, we query the data from the Gaia DR2 catalogue for the region encompassed by this circle. We get stars with a magnitude $G < 17$, since the data from the UCAC4 catalogue does not show stars fainter than $G = 17$. This gives us the data for all the stars in the queried region, so we crossmatch this data with the data from the original study to leave only the stars that are listed in Sampedro's study as being part of the cluster.

@@

Figure \ref{crossmatched_data} shows one example of the plots that were made for each of the clusters, for one of the cases that are confirmed to be a cluster and that will be studied in the next section. Note that, in this case, the proper motion space is pretty compact, showing that the stars are moving together, and there is a clear straight line in the parallax versus magnitude plot, showing that the stars are at the same distance. However, it is interesting to note that NGC 2126 does not look like a cluster when observed on the sky. In the color-magnitude diagram, we can see that some stars have already left the main sequence. We will study this in more detail in the following section.

To filter out the repeated cases due to name inconsistencies, the cluster positions from the study of Cantat-Gaudin were also added to the plots, so in the case of a repeated cluster, we would see it near the center in the position plot, which corresponds to the first plot as shown in Figure \ref{crossmatched_data}.

After examining the plots for all 890 clusters, excluding repetitions, we found that only 14 of them were really compact in proper motion, leading us to believe that the rest of them probably are not actually clusters. These clusters are studied further in the next section.

\section{Estimating the age and interstellar extinction of the cluster}
In this section we discuss the procedure followed to estimate the age and interstellar extinction of some clusters. We applied this method to a total of 14 clusters, some of which are taken from the list of verified clusters in the study of Cantat-Gaudin, and the rest are cases from the study of Sampedro which we have verified in the previous section. The complete list of investigated clusters is given in Table \ref{tab:clusters}, along with some parameters. To obtain the list of probable members, the UPMASK classification method \cite{cantat-gaudin} was applied.
Note that a lower parallax means a greater distance.

\begin{table}[h!]
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Cluster} & \textbf{Position (deg)} & \textbf{Parallax (mas)} & \textbf{Members} \\
\hline
Alessi 3 & $(109.95, -46.05)$ & $3.5619$ & $243$ \\
\hline
Berkeley 100 & $(351.48, 63.78)$ & $0.1461$ & $57$ \\
\hline
Collinder 421 & $(305.83, 41.69)$ & $0.8169$ & $283$ \\
\hline
ESO 332 08 & $(253.80, -40.96)$ & $0.5332$ & $417$ \\
\hline
Herschel 1 & $(116.72, 0.15)$ & $3.3385$ & $149$ \\
\hline
NGC 752 & $(29.24, 37.77)$ & $2.2245$ & $263$ \\
\hline
NGC 1039 & $(40.51, 42.74)$ & $1.9350$ & $659$ \\
\hline
NGC 2126 & $(90.65, 49.91)$ & $0.7490$ & $151$ \\
\hline
NGC 2169 & $(92.12, 13.94)$ & $0.9863$ & $133$ \\
\hline
NGC 2287 & $(101.51, -20.70)$ & $1.3503$ & $673$ \\
\hline
NGC 2360 & $(109.44, -15.63)$ & $0.8994$ & $695$ \\
\hline
NGC 2479 & $(118.76, -17.73)$ & $0.6311$ & $159$ \\
\hline
Ruprecth 8 & $(105.41, -13.56)$ & $0.4278$ & $138$ \\
\hline
Stock 6 & $(35.93, 63.79)$ & $0.9536$ & $144$ \\
\hline
\end{tabular}
\caption{List of clusters that have been studied, including their mean position (given in ($\alpha$, $\delta$)) and mean parallax, computed using the data from Gaia DR2, and the number of member stars of the cluster.}
\label{tab:clusters}
\end{table}

@@

Given that the stars in an open cluster are relatively close together, we shift every star in the plot by the distance modulus of the mean parallax of the cluster. We chose this approach instead of using shifting each star by its individual distance modulus because for more distant stars the error in the parallax is higher, so there might be a lot of discrepancy between the parallaxes of some of the stars in the cluster and the CMD would be less clear. Another option would have been to treat the distance modulus as a third free parameter to estimate via isochrone fitting, which would also avoid this problem but would make the process harder.

An isochrone is the line that would be drawn in a CMD if we varied the mass of the star at a fixed age. Since the stars in an open cluster have the same age, we can find an isochrone that fits the CMD of the stars in the cluster to determine its age. The isochrones used are from PARSEC \cite{parsec}, with the photometric system from Gaia DR2, and including the interstellar extinction $A_V$ on a star-to-star basis \cite{girardi}. The metallicity of the stars, which measures the abundance of heavy elements, is also included in the isochrones used, with a value of $Z = 0.0152$, which is the metallicity of the Sun. An example of the result of fitting an isochrone to the cluster data can be seen in Figure \ref{nice-isochrone}.

@@

\section{Issues when fitting isochrones}
\subsection{Unresolved binary stars}
Many stellar systems are binary stars, systems formed by two stars orbiting each other. If they are distant, some these systems may be unresolved and seen as a single object. In a CMD, unresolved binary stars would appear as a secondary line in the main sequence which is brighter than the rest. A clear example of this can be seen in Figure \ref{nice-isochrone}. This can make the main sequence line in the CMD appear wider than it actually is, and can be a problem when fitting isochrones.

@@

\subsection{Differential extinction}
In this work, the reddening due to the interstellar extinction had a fixed value and was included in the isochrones used to estimate the parameters of the clusters. However, in some cases, like for instance Stock 6, the extinction is not the same for all stars in the cluster. In the case of Stock 6, more distant stars appeared to have significantly more extinction, and this makes the stars appear more scattered in the CMD, making the fitting harder. To avoid this problem, we would have to estimate the extinction for each star individually, and apply the corrections on each of the stars, instead of applying a global extinction for all the stars in the cluster.
